Interoception Scale (Study 2) - Data Cleaning

### 2025-12-29 paper/manuscript.qmd
added omega for CEFSA, PI and PHQ

The Černis Felt Sense of Anomaly scale - short form [CEFSA-14, @vcernis2024measuring] measures 7 dissociative experiences with 14 items on a 5-point Likert scale (0 = *Never*, 4 = *Always*). This scale is designed to measure dissociative experiences in adolescence and adulthood, specifically focusing on the felt sense of anomaly-type dissociation. It includes Anomalous Experience of the Self, Anomalous Experience of the Body, Anomalous Experiences of Emotion, Altered Sense of Familiarity, Altered Sense of Connection, Altered Sense of Agency, and Altered Sense of Reality.

The Primal World Beliefs Inventory [PI-18, @clifton2021brief] measures higher-order beliefs about the basic character of the world: Good, Safe, Enticing, and Alive. The 18 items are presented on a 6-point Likert scale (0 = *Strongly Disagree*, 5 = *Strongly Agree*).

@@

The Single Item Life Satisfaction scale [SILS, @jovanovic2020longer] was followed by the PHQ-4 [@kroenke2009ultra] assessing depression and anxiety symptoms with 4 items. We used the refined version of the PHQ-4 [@makowski2025measuring], which adds an additional response option ("Once or twice"), increasing sensitivity to subclinical fluctuations.

>>>>

The Černis Felt Sense of Anomaly scale - short form [CEFSA-14, @vcernis2024measuring] measures 7 dissociative experiences with 14 items on a 5-point Likert scale (0 = *Never*, 4 = *Always*). This scale is designed to measure dissociative experiences in adolescence and adulthood, specifically focusing on the felt sense of anomaly-type dissociation. It includes Anomalous Experience of the Self (ω= 0.75), Anomalous Experience of the Body (ω= 0.77), Anomalous Experiences of Emotion (ω= 0.81), Altered Sense of Familiarity (ω= 0.72), Altered Sense of Connection (ω= 0.86), Altered Sense of Agency (ω= 0.78), and Altered Sense of Reality (ω= 0.58). The internal consistency of the subscales were mostly good with the exception of Sense of Reality.

The Primal World Beliefs Inventory [PI-18, @clifton2021brief] measures higher-order beliefs about the basic character of the world: Good, Safe, Enticing, and Alive. The 18 items are presented on a 6-point Likert scale (0 = *Strongly Disagree*, 5 = *Strongly Agree*). In this study, the internal consistency of the subscales were all good: ωGood = 0.84, ωSafe = 0.80, ωEnticing = 0.80 and ωAlive = 0.84.

@@

The Single Item Life Satisfaction scale [SILS, @jovanovic2020longer] was followed by the PHQ-4 [@kroenke2009ultra] assessing depression and anxiety symptoms with 4 items. We used the refined version of the PHQ-4 [@makowski2025measuring], which adds an additional response option ("Once or twice"), increasing sensitivity to subclinical fluctuations. PhQ-Anxiety showed excellent reliability (ω= 0.91) and PHQ-Depression showed good reliability (ω= 0.84).

### 2025-12-29 study2/analysis/1_cleaning.qmd
added omega for CEFSA, PI and PHQ

- MAIA [x]
- MINT []
- BPQ [x]
- IAS [x]
- TAS-20 [x]
- B-ERS []
- CERQ-Short []
- CEFSA-14 []
- PI-18 []
- SILS (only one item)
- PHQ-4 []

@@

@@

df <- df |>
  mutate(TAS_DIF_9 = 6 - TAS_DIF_9,
         TAS_DIF_13 = 6- TAS_DIF_13 , 
         TAS_EOT_15 = 6- TAS_EOT_15,
         TAS_EOT_18 = 6- TAS_EOT_18,
         TAS_EOT_19 = 6- TAS_EOT_19)

>>>>

- MAIA [x]
- MINT []
- BPQ [x]
- IAS [x]
- TAS-20 [x]
- B-ERS [x]
- CERQ-Short [x]
- CEFSA-14 [x]
- PI-18 [x]
- SILS (only one item)
- PHQ-4 []

@@

```{r}
cefsa_emo <- df[, grepl("^CEFSA_Emotion_", names(df))]
cefsa_con <- df[, grepl("^CEFSA_Connection_", names(df))]
cefsa_body <- df[, grepl("^CEFSA_Body_", names(df))]
cefsa_rea <- df[, grepl("^CEFSA_Reality_", names(df))]
cefsa_fam <- df[, grepl("^CEFSA_Familiarity_", names(df))]
cefsa_age <- df[, grepl("^CEFSA_Agency_", names(df))]
cefsa_self <- df[, grepl("^CEFSA_Self_", names(df))]

psych::omega(cefsa_emo, nfactors = 1)
psych::omega(cefsa_con, nfactors = 1)
psych::omega(cefsa_body, nfactors = 1)
psych::omega(cefsa_rea, nfactors = 1)
psych::omega(cefsa_fam, nfactors = 1)
psych::omega(cefsa_age, nfactors = 1)
psych::omega(cefsa_self, nfactors = 1)
```

```{r}

df[grepl("^PI18.*_R$", names(df))] <- 6 - df[grepl("^PI18.*_R$", names(df))]  # Reverse

pi_alive <- df[grep("^PI18.*A_", names(df))]
pi_safe <- df[grep("^PI18.*GS", names(df))]
pi_enticing <- df[grep("^PI18.*GE", names(df))]
pi_good <- df[c(grep("^PI18.*GS", names(df)), grep("^PI18.*GE", names(df)))]

psych::omega(pi_alive, nfactors = 1)
psych::omega(pi_safe, nfactors = 1)
psych::omega(pi_enticing, nfactors = 1)
psych::omega(pi_good, nfactors = 1)
```

```{r}
phq_a <- df[grep("^PHQ4_Anxiety", names(df))]
phq_d <- df[grep("^PHQ4_Depression", names(df))]

psych::omega(phq_a, nfactors = 1)
psych::omega(phq_d, nfactors = 1)
```

@@

df <- df |>
  mutate(TAS_DIF_9 = 6 - TAS_DIF_9,
         TAS_DIF_13 = 6- TAS_DIF_13, 
         TAS_EOT_15 = 6- TAS_EOT_15,
         TAS_EOT_18 = 6- TAS_EOT_18,
         TAS_EOT_19 = 6- TAS_EOT_19)
